---
title: "R Basics"
author: "Liang Bo Wang"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: false
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


# Basic Algebra

## Mathematical operations are vectorized

By default, opreations are element-wise and will be **broadcasted** to each pair of inputs.

```{r}
1:5 + 6:10
c(9, 5, 2, 7) * c(5, 5, 6, 6)
689 > 609
1:3 == 3:1
```

## Numerical comparison
Don't use `==` or `!=` to compare numbers, which easily gives round-off errors. Float numbers (e.g. 1.414 and 55.66) have precision limit 

```{r}
sqrt(1:5) ^ 2 == 1:5
all.equal(sqrt(1:5) ^ 2, 1:5)
all.equal(sqrt(2) ^ 2, 3)
isTRUE(all.equal(sqrt(2) ^ 2, 3))
```

## Vectors and Variables

```{r}
x <- 1:5
y <- c(-1, 1, 2, 3:4)
x - y
all(x^2 >= 1)
any(y < 0)
(y > 0) & (y ^ 2 < 3)
```


## Strings and Factors

```{r}
c("Bioinfo", "and", "Biostat", "Core", "Lab")
(gender <- factor(
    c("male", "female", "male", "female", "female")
))
levels(gender)
as.integer(gender)  # convert factor into integers
```


## Vector indexing

```{r}
my_vec <- (1:5) ^ 2
all.equal(seq_along(my_vec), seq_len(length(my_vec)))
names(my_vec)  <- c("one", "four", "nine", "sixteen", "twentyfive")
my_vec[1:2]
my_vec[-c(3, 5)]
my_vec[c("one", "four", "sixteen")]
```

```{r}
my_vec <- (1:10) ^ 2
my_vec[my_vec < 10]
my_vec[my_vec > 15 & my_vec < 30]
```


# Function and Control Flow
Not going to use much today.

```{r}
my_magic_salary_booster <- function(income) {
    loser_level <- income < 22e3
    income[loser_level] <- income[loser_level] * 2
    winner_level <- !loser_level  # loser_level is a logical vector
    income[winner_level] <- income[winner_level] ^ 2
    
    income   # last evaluated value being returned
}

society_minature_income <- c(4e3, 8e3, 15e3, 5e4, 10e4)
names(society_minature_income) <- c("arbeit", "master", "NSC", "professor", "boss")
my_magic_salary_booster(society_minature_income)
```


# List and Data Frame

## List
```{r}
(mylist <- list(a = 1:10, b = c("a", "b", "c"), 1:10 < 4))
mylist[[1]]
mylist$a
names(mylist)
```

## Data Frame
```{r}
my_calendar <- data.frame(
    month = month.abb, 
    date = c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31),
    stringsAsFactors = FALSE
)
my_calendar$month %in% c("Feb", "May")
my_calendar[1:2, ]
my_calendar[my_calendar$date > 30, c("month")]
```

To access the general info about a data frame

```{r}
row.names(my_calendar)
colnames(my_calendar)
dim(my_calendar)
c(nrow(my_calendar), ncol(my_calendar))
```


`with` expose col names of data frame as variables

```{r}
with(my_calendar, date[month %in% c("May", "Jun", "Aug")])
```

There are many built-in data frames.

```{r}
head(iris)
iris[iris$Sepal.Width > 4, ]
```




